---
title: "'ggplot' 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```


## create the weather data

```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


## making new plots

start with old plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

add labels: 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp(C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  )
```

x axis tck marks etc

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp(C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20), 
    labels = c("-15C", "-5C", "20C")
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
# scale_y_sqrt() has the same function
```

A `sqrt` transformation is fine here. But some of the negative values of y are dropped from
plotting because the you cannot square root a negative value. 


## colors

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp(C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) + 
  scale_color_hue(
    name = "Weather Station", 
    h = c(100, 300)
  )
```

use viridis packages

```{r}
ggp_base = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp(C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base
```


## themes

```{r}
ggp_base + 
  theme(legend.position = "bottom") +
  theme_bw()
```

at here, the order of command matter. we do legend first, then do `theme_minimal`, it gonna
remove everything. 

`theme_bw()` can change change the color of the backgroud. grey background and white line to 
white background and black lines.
`theme_minimal()` remove the line frame of the plots
`theme_classic` gives a classical plot









