---
title: "'ggplot' 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```


## create the weather data

```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```

## advanced scatter plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(~ name)
```

Also I want to learn about precipitation, so let’s do that.

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```


The previous notes were lost, starts from here!

## Some extra stuff

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_smooth(size = 2, se = FALSE)
```
you can pick whatever you think that is useful to show up.

2d density

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_hex() +
  facet_grid(~name)
```
within this area, how much data points exist in there


## more kinds of plots!!

understand the distribution of a single variable
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, color = name)) +
  geom_histogram()
```

we don't like those bars stack with each other, so we separate those bars

Also, the color is outside the bar and we don't like that, then we use
`fill = name`

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")
```

or we have 3 separate panel

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(~name)
```

now let's try density plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

boxplots...

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

violin plot...

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

ridge plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```


## saving a plot

```{r}
ggp_ridge_temp = weather_df %>% ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_ridge_temp)
```





